---
title: "Popular names"
subtitle: "Week 6 Activity"
author: "Branson T Fausey"
date: "Sunday, 6 October 2024"
output: 
  html_notebook:
    fig_height: 3
    fig_width: 5
---

# Set Up

```{r message=FALSE}
# load necessary packages
library(tidyverse)
library(dcData)
library(mosaic)
library(esquisse)
```


```{r}
# Load the BabyNames data into RStudio environment (from dcData package)
data("BabyNames")

```

# Activity

**Goal**: By the end of the activity, each team will create a graph (like Figure A.1 on p 159) showing ups and downs in the popularity of the names at your table according to the U.S. social security registry.  

___

### Step 1


```{r}
## Helpful to look at the data a few different ways (see Ch 9 for more).
help(BabyNames)
BabyNames %>%
  head(5)
nrow(BabyNames)
```
The data frame consists of 4 variables across 1.79 million cases. Each case represents a name, given the gender and year.

___

### Step 2

[This should be done by hand on paper, nothing to show in the HTML doc]

___

### Step 3 

##### What variable(s) from the raw data table do not appear at all in the graph?

The 'sex' variable does not appear at all in the target graph, and 'count' appears to be transformed.

##### What variable(s) in the graph are similar to corresponding variables in the raw data table, but might have been transformed in some way?

The 'Popularity' appears to be the 'count' variable of both male and females combined

___

### Step 4

##### Have cases been filtered out?

All cases have been filtered out except for the desired names.

##### Have cases been grouped and summarized within groups in any way?

Cases have been grouped by 'name' and 'year', removing the 'sex' variable.

##### Have any new variables been introduced?  If so, what the relationship between the new variables and existing variables?

It appears the only new variable is 'Popularity', which simply appears to be the 'count' of both males and females.

___

### Step 5

- Group the cases by 'name' and 'year'.
- Filter out all the unwanted names.
- Summarise the 'count' variable to a sex-combined 'total' variable.

___

### Step 6

[This should be done by hand on paper, nothing to show in the HTML doc]



### Step 7

```{r}
GRdata <- BabyNames %>%
group_by(name, year) %>%
filter(name %in% c("Branson", "Emma", "Neil", "Susan")) %>%
summarise(total = sum(count))
```


### Graph It!

Here's the sample code provided by the book:

```
GlyphReadyForm %>%
  ggplot(aes(x = year, y = total, group = name)) + 
  geom_line(size = 1, alpha = 0.5, aes(color = name)) + 
  ylab("Popularity") + xlab("Year")
```
```{r}
ggplot(GRdata) +
    aes(x = year, y = total, colour = name) +
    geom_step(linewidth = 1) +
    scale_color_hue(direction = 1) +
    labs(
        x = "Year",
        y = "Popularity",
        title = "Popularity of Names",
        color = "Names:"
    ) +
    theme_minimal()
```


Note: if someone in your group has a name that hasn't been registered with the US Social Security Administration for a baby born in the United States, it's okay that the student's name may not appear on the plot.